Secure Composition \\ of Robust and Optimising Compilers

### 2024-05-20 csf2/main.tex
implement some todos

\usepackage[colorinlistoftodos]{todonotes}
\usepackage[inference]{semantic}
\usepackage[switch]{lineno}
\usepackage{fontawesome5}
\usepackage{listofitems}
\usepackage{breakcites}
\usepackage{glossaries}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{stmaryrd}
\usepackage{marvosym}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{nameref}
\usepackage{amsthm}
\usepackage{xspace}
\usepackage{xfrac}
\usepackage{tikz}
\usepackage{soul}
\usepackage{bm}

@@

\[
  \begin{array}{rcll}
    \event[^b_{ct}] &:=& \\
    \event[_{ct}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n} \mid \emptyevent \mid \lightning \\ 
  \end{array}
\]

@@

The combination of \gls*{ms} and \gls*{scct} is, intuitively, the intersection of these properties, \gls*{msscct}.
However, \gls*{ms} uses a simpler trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
To remedy this issue, let $\sim_{msscct}$ be a cross-language trace relation that relates the basic memory safety trace model with the memory safety trace model by ignoring security tags and replacing the new pre-events with the empty event:
\[
  \typerulenolabel{mscctrel:drop:tag}{}{\event[_b];\sim_{msscct}\event[_b];\securitytag}
\]
\[
  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{msscct}\ev{Branch n};\securitytag}
  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{msscct}\ev{Binop n};\securitytag}
\]

@@

@@

This section defines programming languages that the secure compilers defined in the next section will use.
To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, and $\Lscct$ which share many common elements (presented in \Cref{subsec:cs:defs}).
$\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
However, not all $\Ltms$ programs are \gls*{sms}.
That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
Language $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls*{ms} (\Cref{subsec:lsms}).
$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{cct} to hold.

@@

>>>>

\usepackage[colorinlistoftodos]{todonotes}
\usepackage[inference]{semantic}
\usepackage[switch]{lineno}
\usepackage{halloweenmath}
\usepackage{fontawesome5}
\usepackage{listofitems}
\usepackage{breakcites}
\usepackage{glossaries}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{stmaryrd}
\usepackage{marvosym}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{nameref}
\usepackage{amsthm}
\usepackage{xspace}
\usepackage{xfrac}
\usepackage{tikz}
\usepackage{soul}
\usepackage{bm}

@@

\[
  \begin{array}{rrcl}
    (\text{Pre-Event}) & \event[^b_{ct}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n}\\
                       &&\mid& \ev{Branch\ }n \mid \ev{Binop\ }n\\
    (\text{Security Tag}) & \securitytag &:=& \lock \mid \unlock\\ 
    (\text{Event}) & \event[_{ct}] &:=& \event[^b_{ct}];\securitytag \mid \emptyevent \mid \lightning \\ 
  \end{array}
\]

@@

The combination of \gls*{ms} and \gls*{scct} is, intuitively, the intersection of these properties, \gls*{msscct}.
However, \gls*{ms} uses a simpler trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
To remedy this issue, let $\sim_{msscct}$ be a cross-language trace relation that relates the basic memory safety trace model with the memory safety trace model by ignoring security tags and replacing the new pre-events with the empty event:
\[
  \typerulenolabel{mscctrel:drop:tag}{}{\event[_{ms}]\sim_{msscct}\event[^b_{ct}];\securitytag}
\]
\[
  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{msscct}\ev{Branch n};\securitytag}
  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{msscct}\ev{Binop n};\securitytag}
\]

@@

\[
  \begin{array}{rrcl}
    (\text{Pre-Event}) & \event[^b_{\mathghost}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n}\\
                       &&\mid& \ev{Branch\ }n \mid \ev{Binop\ }n \mid \ev{Spec} \mid \ev{Rlb}\\
    (\text{Spectre Variant}) & vX &:=& \operatorname{PHT} \\
    (\text{Security Tag}) & \securitytag &:=& \lock_{vX} \mid \unlock\\ 
    (\text{Event}) & \event[_{\mathghost}] &:=& \event[^b_{\mathghost}];\securitytag \mid \emptyevent \mid \lightning \\ 
  \end{array}
\]

@@

This section defines programming languages that the secure compilers defined in the next section will use.
To this end, this section defines the languages $\Ltms$, $\Ltrg$, $\Lms$, $\Lscct$, and $\Lspec$ which share many common elements (presented in \Cref{subsec:cs:defs}).
$\Ltms$ is the only statically typed language and exhibits the property that all well-typed programs are \gls*{tms} (\Cref{subsec:ltms}).
However, not all $\Ltms$ programs are \gls*{sms}.
That is, there are well-typed $\Ltms$ programs that perform an out-of-bounds access.
Language $\Ltrg$ is untyped and does not provide any guarantees with regards to \gls*{ms} (\Cref{subsec:lsms}).
$\Lms$ is exactly the same language as $\Ltrg$, but this paper still distinguishes the two for sake of readability (\Cref{subsec:lms}).
All three languages --- so $\Ltms$, $\Ltrg$, and $\Lms$ --- assume \gls*{cct} to hold.

@@

Lastly, modern processors also employ speculative execution to achieve speedups, but fail to rollback properly, so that data may be left over at a cache, so that an attacker can exploit the difference in runtime for cache-hits versus cache-misses. 
$\Lspec$ extends $\Lscct$ to model this situation, while all previous languages trivially satisfy \gls*{ss}, since they do not support speculative execution at all.
